---
title: Compute Gini Inequality Coefficient Given Data Vector (One Variable)
description: |
  View a vector positive vectors as population, calculate discrete GINI inequality measure.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compute Gini Inequality Coefficient Given Sample (One Variable)}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
urlcolor: blue
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

View a vector positive vectors as population, calculate discrete GINI inequality measure. This file works out how the [ff_dist_gini_vector_pos](https://fanwangecon.github.io/REconTools/reference/ff_dist_gini_vector_pos.html) function works from [Fan](https://fanwangecon.github.io/)'s *[REconTools](https://fanwangecon.github.io/REconTools/)* Package. See also the [fs_gini_disc](https://fanwangecon.github.io/R4Econ/math/func_ineq/htmlpdfr/fs_gini_disc.html) from *[R4Econ](https://fanwangecon.github.io/R4Econ/)*. See also the [ff_dist_gini_random_var](https://fanwangecon.github.io/REconTools/reference/ff_dist_gini_random_var.html) function for GINI for discrete random variables.

There is an vector values (all positive). This could be height information for N individuals. It could also be income information for N individuals. Calculate the [GINI](https://en.wikipedia.org/wiki/Gini_coefficient) coefficient treating the given vector as population. This is not an estimation exercise where we want to estimate population gini based on a sample. The given array is the population. The population is discrete, and only has these N individuals in the length n vector.

See the formula below, note that when the sample size is small, there is a limit to inequality using the formula defined below given each $N$. So for small $N$, can not really compare inequality across arrays with different $N$, can only compare arrays with the same $N$.

## Formula

Given monotonimcally increasing array $X$, with $x_1,...,x_N$:

1. There is a box, width = 1, height = 1
2. The width is discretized into $N$ individuals, so each individual's width is $\frac{1}{N}$
3. The height is normalized to 1, for the nth individual, total height is the sum of all $x$, so need to rescale all bars by $\sum_i^{N} x_i$

The GINI formula used here is:
$$
 GINI =
  1 - \frac{2}{N+1}
  \cdot
  \left(\sum_{i=1}^N \sum_{j=1}^{i} x_j\right)
  \cdot
  \left(
    \sum_{i=1}^N x_i
  \right)^{-1}
$$

Derive the formula in the steps below.

### Step 1 Area Formula

$$
 \Gamma = \sum_{i=1}^N \frac{1}{N} \cdot \left(
     \sum_{j=1}^{i} \left(
      \frac{x_j}{\sum_{\widehat{j}=1}^N x_{\widehat{j}} }
      \right)
  \right)
$$

### Step 2 Total Area Given Perfect equality

With perfect equality $x_i=a$ for all $i$, so need to divide by that.

$$
 \Gamma^{\text{equal}} = \sum_{i=1}^N \frac{1}{N} \cdot \left(
     \sum_{j=1}^{i} \left(
      \frac{a}{\sum_{\widehat{j}=1}^N a }
      \right)
  \right)
  = \frac{N+1}{N}\cdot\frac{1}{2}
$$

As the number of elements of the vecotr increases:
$$
 \lim_{N \rightarrow \infty}\Gamma^{\text{equal}}
  = \lim_{N \rightarrow \infty} \frac{N+1}{N}\cdot\frac{1}{2}
  = \frac{1}{2}
$$

### Step 3 Arriving at Finite Vector Gini Formula

Given what we have from above, we obtain the gini formula, divide by total area below 45 degree line.

$$
 GINI =
  1 -
  \left(\sum_{i=1}^N \sum_{j=1}^{i} x_j\right)
  \cdot
  \left(
    N \cdot \sum_{i=1}^N x_i
  \right)^{-1}
  \cdot
  \left( \frac{N+1}{N}\cdot\frac{1}{2} \right)^{-1}
  =
  1 - \frac{2}{N+1}
  \cdot
  \left(\sum_{i=1}^N \sum_{j=1}^{i} x_j\right)
  \cdot
  \left(
    \sum_{i=1}^N x_i
  \right)^{-1}
$$

### Step 4 Maximum Inequality given N

Suppose $x_i=0$ for all $i<N$, then:

$$
 GINI^{x_i = 0 \text{ except } i=N}
 =
  1 - \frac{2}{N+1}
  \cdot
  X_N
  \cdot
  \left(
    X_N
  \right)^{-1}
 =
 1 - \frac{2}{N+1}
$$

$$
 \lim_{N \rightarrow \infty} GINI^{x_i = 0 \text{ except } i=N}
 =
 1 -
 \lim_{N \rightarrow \infty}
 \frac{2}{N+1}
 = 1
$$

Note that for small N, for example if $N=10$, even when one person holds all income, all others have 0 income, the formula will not produce gini is zero, but that gini is equal to $\frac{2}{11}\approx 0.1818$. If $N=2$, inequality is at most, $\frac{2}{3}\approx 0.667$.

$$
 MostUnequalGINI\left(N\right) = 1 - \frac{2}{N+1} = \frac{N-1}{N+1}
$$

## Implement GINI Formula in R

The **GINI** formula just derived is trivial to compute.

1. scalar: $\frac{2}{N+1}$
2. cumsum: $\sum_{j=1}^{i} x_j$
3. sum of cumsum: $\left(\sum_{i=1}^N \sum_{j=1}^{i} x_j\right)$
3. sum: $\sum_{i=1}^N X_i$

There are no package dependencies. Define the formula here:

```{r}
# Load Library
rm(list = ls())
# Formula, directly implement the GINI formula Following Step 4 above
fv_dist_gini_vector_pos_test <- function(ar_pos) {
  # Check length and given warning
  it_n <- length(ar_pos)
  if (it_n <= 100)  warning('Data vector has n=',it_n,', max-inequality/max-gini=',(it_n-1)/(it_n + 1))
  # Sort
  ar_pos <- sort(ar_pos)
  # formula implement
  fl_gini <- 1 - ((2/(it_n+1)) * sum(cumsum(ar_pos))*(sum(ar_pos))^(-1))
  return(fl_gini)
}
```

### Testing

Generate a number of examples Arrays for testing

```{r}
# Example Arrays of data
ar_equal_n1 = c(1)
ar_ineql_n1 = c(100)

ar_equal_n2 = c(1,1)
ar_ineql_alittle_n2 = c(1,2)
ar_ineql_somewht_n2 = c(1,2^3)
ar_ineql_alotine_n2 = c(1,2^5)
ar_ineql_veryvry_n2 = c(1,2^8)
ar_ineql_mostmst_n2 = c(1,2^13)

ar_equal_n10 = c(2,2,2,2,2,2, 2, 2, 2, 2)
ar_ineql_some_n10 = c(1,2,3,5,8,13,21,34,55,89)
ar_ineql_very_n10 = c(1,2^2,3^2,5^2,8^2,13^2,21^2,34^2,55^2,89^2)
ar_ineql_extr_n10 = c(1,2^2,3^3,5^4,8^5,13^6,21^7,34^8,55^9,89^10)

# Uniform draw testing
ar_unif_n1000 = runif(1000, min=0, max=1)

# Normal draw testing
ar_norm_lowsd_n1000 = rnorm(1000, mean=100, sd =1)
ar_norm_lowsd_n1000[ar_norm_lowsd_n1000<0] = 0
ar_norm_highsd_n1000 = rnorm(1000, mean=100, sd =20)
ar_norm_highsd_n1000[ar_norm_highsd_n1000<0] = 0

# Beta draw testing
ar_beta_mostrich_n1000 = rbeta(1000, 5, 1)
ar_beta_mostpoor_n1000 = rbeta(1000, 1, 5)
ar_beta_manyrichmanypoor_nomiddle_n1000 = rbeta(1000, 0.5, 0.5)
```

Now test the example arrays above using the function based no our formula:

```{r, eval=TRUE, echo=FALSE}
# Hard-Code Small N Dist Tests
cat('\nSmall N=1 Hard-Code\n')
cat('ar_equal_n1:', fv_dist_gini_vector_pos_test(ar_equal_n1), '\n')
cat('ar_ineql_n1:', fv_dist_gini_vector_pos_test(ar_ineql_n1), '\n')

cat('\nSmall N=2 Hard-Code, converge to 1/3, see formula above\n')
cat('ar_ineql_alittle_n2:', fv_dist_gini_vector_pos_test(ar_ineql_alittle_n2), '\n')
cat('ar_ineql_somewht_n2:', fv_dist_gini_vector_pos_test(ar_ineql_somewht_n2), '\n')
cat('ar_ineql_alotine_n2:', fv_dist_gini_vector_pos_test(ar_ineql_alotine_n2), '\n')
cat('ar_ineql_veryvry_n2:', fv_dist_gini_vector_pos_test(ar_ineql_veryvry_n2), '\n')

cat('\nSmall N=10 Hard-Code, convege to 9/11=0.8181, see formula above\n')
cat('ar_equal_n10:', fv_dist_gini_vector_pos_test(ar_equal_n10), '\n')
cat('ar_ineql_some_n10:', fv_dist_gini_vector_pos_test(ar_ineql_some_n10), '\n')
cat('ar_ineql_very_n10:', fv_dist_gini_vector_pos_test(ar_ineql_very_n10), '\n')
cat('ar_ineql_extr_n10:', fv_dist_gini_vector_pos_test(ar_ineql_extr_n10), '\n')

# Uniform Dist Tests
cat('\nUNIFORM Distribution\n')
cat('ar_unif_n1000:', fv_dist_gini_vector_pos_test(ar_unif_n1000), '\n')

# Normal Dist Tests
cat('\nNORMAL Distribution\n')
cat('ar_norm_lowsd_n1000:', fv_dist_gini_vector_pos_test(ar_norm_lowsd_n1000), '\n')
cat('ar_norm_highsd_n1000:', fv_dist_gini_vector_pos_test(ar_norm_highsd_n1000), '\n')

# Beta Dist Tests
cat('\nBETA Distribution\n')
cat('ar_beta_mostpoor_n1000:', fv_dist_gini_vector_pos_test(ar_beta_mostpoor_n1000), '\n')
cat('ar_beta_manyrichmanypoor_nomiddle_n1000:', fv_dist_gini_vector_pos_test(ar_beta_manyrichmanypoor_nomiddle_n1000), '\n')
cat('ar_beta_mostrich_n1000:', fv_dist_gini_vector_pos_test(ar_beta_mostrich_n1000), '\n')

# Example when it does not work
ar_unif_n1000_NEGATIVE = runif(1000, min=-1, max=1)
cat('\n\nSHOULD/DOES NOT WORK TEST\n ar_unif_n1000_NEGATIVE:', fv_dist_gini_vector_pos_test(ar_unif_n1000_NEGATIVE), '\n')
```
