---
title: OPTI BISECT--Solve Roots for N (Rows) Strictly Monotonic Functions with Single
  Zero-Crossing
urlcolor: blue
vignette: |
  %\VignetteIndexEntry{OPTI BISECT--Solve Roots for N (Rows) Strictly Monotonic Functions with Single Zero-Crossing} %\VignetteEngine{knitr::rmarkdown} \usepackage[utf8]{inputenc}
---

Back to **[Fan](https://fanwangecon.github.io/)**'s REconTools **[Table of Content](https://fanwangecon.github.io/REconTools/)**. Also see [R4Econ](https://fanwangecon.github.io/R4Econ/) for R examples and [Intro Stats with R](https://fanwangecon.github.io/Stat4Econ/) for R introductions.

# Objective

This file works out how the [ff_opti_bisect_pmap_multi](https://fanwangecon.github.io/REconTools/reference/ff_opti_bisect_pmap_multi.html) function works from [Fan](https://fanwangecon.github.io/)'s *[REconTools](https://fanwangecon.github.io/REconTools/)* Package. Also see a closely related file from [R4Econ](https://fanwangecon.github.io/R4Econ/), [DPLYR Bisection](https://fanwangecon.github.io/R4Econ/optimization/root_bisect/fs_func_graph_eval.html), which works out in greater detail the bisection algorithm.

We want solve for function $0=f(z_{ij}, x_i, y_i, \textbf{X}, \textbf{Y}, c, d)$. There are $i$ functions that have $i$ specific $x$ and $y$. For each $i$ function, we evaluate along a grid of feasible values for $z$, over $j\in J$ grid points, potentially looking for the $j$ that is closest to the root. $j$ could be interpreted as iterations. $\textbf{X}$ and $\textbf{Y}$ are arrays common across the $i$ equations, and $c$ and $d$ are constants.

As we iterate over $j$, find the $z_{ij}$ that is $i$ specific that allows for $f_{i}(z_{ij})=0$.

In the code below, we develop the bisection code line by line, then we write it as a function using pmap. At the end, we test the concurrent bisection code over tibble dataframe with a set of straight lines (negative intercepts and positive slopes), and the log-linear solution for optimal allocation function from [fan](http://fanwangecon.github.io/)'s [PrjOptiAlloc](https://fanwangecon.github.io/PrjOptiAlloc/) Project.

## Bisection Algorithm

This is how I implement the bisection algorithm, when we know the bounding minimum and maximum to be below and above zero already.

1. Evaluate $f^0_a = f(a^0)$ and $f^0_b = f(b^0)$, min and max points.
2. Evaluate at $f^0_p = f(p^0)$, where $p_0 = \frac{a^0+b^0}{2}$.
3. if $f^i_a \cdot f^i_p < 0$, then $b_{i+1} = p_i$, else, $a_{i+1} = p_i$ and $f^{i+1}_a = p_i$.
4. iteratre until convergence.

And for the function we want to bisect over, as stated earlier, it has three types of variables:

1. variables in file for grouping, each group is an individual for whom we want to calculate optimal choice for using bisection.
2. variables that change each row and each iteration, the root we are looking for.
3. scalar and array values that are applied to every row and persist across iterations.

# Set up for Implementation

Invoking the algorithm uses several key tidyverse functions:

1. *pmap* from [purrr](https://purrr.tidyverse.org/index.html) to evaluate function row by row.
2. *case_when* from [dplyr](https://dplyr.tidyverse.org/reference/case_when.html) to set new lower and upper bound.
3. *pivot_wider* and *pivot_longer* from [tidyr](https://tidyr.tidyverse.org/index.html) to reshape resulting file if record all iteration results.

## Load Packages

```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
library(tibble)
library(tidyr)
library(purrr)
library(dplyr)
library(knitr)
library(ggplot2)
library(kableExtra)
```

## Set up Input Arrays

There is a function that takes $M=Q+P$ inputs, we want to evaluate this function $N$ times. Each time, there are $M$ inputs, where all but $Q$ of the $M$ inputs, meaning $P$ of the $M$ inputs, are the same. In particular, $P=Q*N$.

$$M = Q+P = Q + Q*N$$

Now we need to expand this by the number of choice grid. Each row, representing one equation, is expanded by the number of choice grids. We are graphically searching, or rather brute force searching, which means if we have 100 individuals, we want to plot out the nonlinear equation for each of these lines, and show graphically where each line crosses zero. We achieve this, by evaluating the equation for each of the 100 individuals along a grid of feasible choices.

In this problem here, the feasible choices are shared across individuals.

```{r setup}
# Parameters
fl_rho = 0.20
svr_id_var = 'INDI_ID'

# it_child_count = N, the number of children
it_N_child_cnt = 9
# it_heter_param = Q, number of parameters that are heterogeneous across children
it_Q_hetpa_cnt = 2

# P fixed parameters, nN is N dimensional, nP is P dimensional
ar_nN_A = seq(-2, 2, length.out = it_N_child_cnt)
ar_nN_alpha = seq(0.1, 0.9, length.out = it_N_child_cnt)
ar_nP_A_alpha = c(ar_nN_A, ar_nN_alpha)

# N by Q varying parameters
mt_nN_by_nQ_A_alpha = cbind(ar_nN_A, ar_nN_alpha)

# Choice Grid for nutritional feasible choices for each
fl_N_agg = 100
fl_N_min = 0

# Mesh Expand
tb_states_choices <- as_tibble(mt_nN_by_nQ_A_alpha) %>% rowid_to_column(var=svr_id_var)
ar_st_col_names = c(svr_id_var,'fl_A', 'fl_alpha')
tb_states_choices <- tb_states_choices %>% rename_all(~c(ar_st_col_names))

# display
summary(tb_states_choices)
kable(tb_states_choices) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

## Create Nonlinear Root Function 

These is a test functions over which we will searh for roots. 

```{r}
# Define Implicit Function
ffi_nonlin_dplyrdo <- function(fl_A, fl_alpha, fl_N, ar_A, ar_alpha, fl_N_agg, fl_rho){
  # scalar value that are row-specific, in dataframe already: *fl_A*, *fl_alpha*, *fl_N*
  # array and scalars not in dataframe, common all rows: *ar_A*, *ar_alpha*, *fl_N_agg*, *fl_rho*

  # Test Parameters
  # ar_A = ar_nN_A
  # ar_alpha = ar_nN_alpha
  # fl_N = 100
  # fl_rho = -1
  # fl_N_q = 10

  # Apply Function
  ar_p1_s1 = exp((fl_A - ar_A)*fl_rho)
  ar_p1_s2 = (fl_alpha/ar_alpha)
  ar_p1_s3 = (1/(ar_alpha*fl_rho - 1))
  ar_p1 = (ar_p1_s1*ar_p1_s2)^ar_p1_s3
  ar_p2 = fl_N^((fl_alpha*fl_rho-1)/(ar_alpha*fl_rho-1))
  ar_overall = ar_p1*ar_p2
  fl_overall = fl_N_agg - sum(ar_overall)

  return(fl_overall)
}
```

# Step by Step Implmentation of Algorithm

Generate New columns of a and b as we iteratre, do not need to store p, p is temporary. Evaluate the function below which we have already tested, but now, in the dataframe before generating all permutations, *tb_states_choices*, now the *fl_N* element will be changing with each iteration, it will be row specific. *fl_N* are first min and max, then each subsequent ps.

## Initialize Matrix

First, initialize the matrix with $a_0$ and $b_0$, the initial min and max points:
```{r}

# common prefix to make reshaping easier
st_bisec_prefix <- 'bisec_'
svr_a_lst <- paste0(st_bisec_prefix, 'a_0')
svr_b_lst <- paste0(st_bisec_prefix, 'b_0')
svr_fa_lst <- paste0(st_bisec_prefix, 'fa_0')
svr_fb_lst <- paste0(st_bisec_prefix, 'fb_0')

# Add initial a and b
tb_states_choices_bisec <- tb_states_choices %>%
                            mutate(!!sym(svr_a_lst) := fl_N_min, !!sym(svr_b_lst) := fl_N_agg)

# Evaluate function f(a_0) and f(b_0)
tb_states_choices_bisec <- tb_states_choices_bisec %>% rowwise() %>%
                            mutate(!!sym(svr_fa_lst) := ffi_nonlin_dplyrdo(fl_A, fl_alpha, !!sym(svr_a_lst),
                                                                          ar_nN_A, ar_nN_alpha,
                                                                          fl_N_agg, fl_rho),
                                   !!sym(svr_fb_lst) := ffi_nonlin_dplyrdo(fl_A, fl_alpha, !!sym(svr_b_lst),
                                                                          ar_nN_A, ar_nN_alpha,
                                                                          fl_N_agg, fl_rho))
# Summarize
dim(tb_states_choices_bisec)
summary(tb_states_choices_bisec)
```

## Iterate and Solve for f(p), update f(a) and f(b)

Implement the DPLYR based Concurrent bisection algorithm.

```{r}

# fl_tol = float tolerance criteria
# it_tol = number of interations to allow at most
fl_tol <- 10^-2
it_tol <- 100

# fl_p_dist2zr = distance to zero to initalize
fl_p_dist2zr <- 1000
it_cur <- 0
while (it_cur <= it_tol && fl_p_dist2zr >= fl_tol ) {

  it_cur <- it_cur + 1

  # New Variables
  svr_a_cur <- paste0(st_bisec_prefix, 'a_', it_cur)
  svr_b_cur <- paste0(st_bisec_prefix, 'b_', it_cur)
  svr_fa_cur <- paste0(st_bisec_prefix, 'fa_', it_cur)
  svr_fb_cur <- paste0(st_bisec_prefix, 'fb_', it_cur)

  # Evaluate function f(a_0) and f(b_0)
  # 1. generate p
  # 2. generate f_p
  # 3. generate f_p*f_a
  tb_states_choices_bisec <- tb_states_choices_bisec %>% rowwise() %>%
                              mutate(p = ((!!sym(svr_a_lst) + !!sym(svr_b_lst))/2)) %>%
                              mutate(f_p = ffi_nonlin_dplyrdo(fl_A, fl_alpha, p,
                                                              ar_nN_A, ar_nN_alpha,
                                                              fl_N_agg, fl_rho)) %>%
                              mutate(f_p_t_f_a = f_p*!!sym(svr_fa_lst))
  # fl_p_dist2zr = sum(abs(p))
  fl_p_dist2zr <- mean(abs(tb_states_choices_bisec %>% pull(f_p)))

  # Update a and b
  tb_states_choices_bisec <- tb_states_choices_bisec %>%
                              mutate(!!sym(svr_a_cur) :=
                                       case_when(f_p_t_f_a < 0 ~ !!sym(svr_a_lst),
                                                 TRUE ~ p)) %>%
                              mutate(!!sym(svr_b_cur) :=
                                       case_when(f_p_t_f_a < 0 ~ p,
                                                 TRUE ~ !!sym(svr_b_lst)))
  # Update f(a) and f(b)
  tb_states_choices_bisec <- tb_states_choices_bisec %>%
                              mutate(!!sym(svr_fa_cur) :=
                                       case_when(f_p_t_f_a < 0 ~ !!sym(svr_fa_lst),
                                                 TRUE ~ f_p)) %>%
                              mutate(!!sym(svr_fb_cur) :=
                                       case_when(f_p_t_f_a < 0 ~ f_p,
                                                 TRUE ~ !!sym(svr_fb_lst)))
  # Save from last
  svr_a_lst <- svr_a_cur
  svr_b_lst <- svr_b_cur
  svr_fa_lst <- svr_fa_cur
  svr_fb_lst <- svr_fb_cur

  # Summar current round
  print(paste0('it_cur:', it_cur, ', fl_p_dist2zr:', fl_p_dist2zr))
  summary(tb_states_choices_bisec %>% select(one_of(svr_a_cur, svr_b_cur, svr_fa_cur, svr_fb_cur)))
}
```

## Reshape Wide to long to Wide

To view results easily, how iterations improved to help us find the roots, convert table from wide to long. Pivot twice. This allows us to easily graph out how bisection is working out iterationby iteration.

Here, we will first show what the raw table looks like, the wide only table, and then show the long version, and finally the version that is medium wide.

```{r reshape solution from wide to very long}
# New variables
svr_bisect_iter <- 'biseciter'
svr_abfafb_long_name <- 'varname'
svr_number_col <- 'value'
svr_id_bisect_iter <- paste0(svr_id_var, '_bisect_ier')

# Pivot wide to very long
tb_states_choices_bisec_long <- tb_states_choices_bisec %>%
  pivot_longer(
    cols = starts_with(st_bisec_prefix),
    names_to = c(svr_abfafb_long_name, svr_bisect_iter),
    names_pattern = paste0(st_bisec_prefix, "(.*)_(.*)"),
    values_to = svr_number_col
  )

# Print
summary(tb_states_choices_bisec_long)
head(tb_states_choices_bisec_long %>% select(-one_of('p','f_p','f_p_t_f_a')), 30)
tail(tb_states_choices_bisec_long %>% select(-one_of('p','f_p','f_p_t_f_a')), 30)

# Pivot wide to very long to a little wide
tb_states_choices_bisec_wider <- tb_states_choices_bisec_long %>%
  pivot_wider(
    names_from = !!sym(svr_abfafb_long_name),
    values_from = svr_number_col
  )

# Print
summary(tb_states_choices_bisec_wider)
head(tb_states_choices_bisec_wider %>% select(-one_of('p','f_p','f_p_t_f_a')), 30)
tail(tb_states_choices_bisec_wider %>% select(-one_of('p','f_p','f_p_t_f_a')), 30)
```

## Graph Bisection Iteration Results

Actually we want to graph based on the long results, not the wider. Wider easier to view in table.

```{r reshape solution for graphing}
# Graph results
lineplot <- tb_states_choices_bisec_long %>%
    mutate(!!sym(svr_bisect_iter) := as.numeric(!!sym(svr_bisect_iter))) %>%
    filter(!!sym(svr_abfafb_long_name) %in% c('a', 'b')) %>%
    ggplot(aes(x=!!sym(svr_bisect_iter), y=!!sym(svr_number_col),
               colour=!!sym(svr_abfafb_long_name),
               linetype=!!sym(svr_abfafb_long_name),
               shape=!!sym(svr_abfafb_long_name))) +
        facet_wrap( ~ INDI_ID) +
        geom_line() +
        geom_point() +    
        labs(title = 'Bisection Iteration over individuals Until Convergence',
             x = 'Bisection Iteration',
             y = 'a (left side point) and b (right side point) values',
             caption = 'DPLYR concurrent bisection nonlinear multple individuals') +
      theme(axis.text.x = element_text(angle = 90, hjust = 1))
print(lineplot)
```

# Implement Bisection as A Function over Dataframe

## Define Bisection Algorithm Function 

Note the *df* input must have the not hard-coded input variables in *fc_withroot*, identical names. In particular, *svr_bisect_solu* should specify the name of the root variable: the x variable, which we are changing to solve for zero, whatever it is called in the *df* dataframe.


Inputs:

- *ls_svr_df_in_func*: string list of variables in df that are inputs for fc_withroot.
- *svr_root_x*: the x variable name that appears n fc_withroot. 

```{r define function}
fv_opti_bisect_pmap_multi_test <- function(df, fc_withroot,
                                           fl_lower_x, fl_upper_x,
                                           ls_svr_df_in_func,
                                           svr_root_x = 'x',
                                           it_iter_tol = 50, fl_zero_tol = 10^-5,
                                           bl_keep_iter = TRUE,
                                           st_bisec_prefix = 'bisec_',
                                           st_lower_x = 'a', st_lower_fx = 'fa',
                                           st_upper_x = 'b', st_upper_fx = 'fb') {

  
# A. common prefix to make reshaping easier
svr_a_lst <- paste0(st_bisec_prefix, st_lower_x, '_0')
svr_b_lst <- paste0(st_bisec_prefix, st_upper_x, '_0')
svr_fa_lst <- paste0(st_bisec_prefix, st_lower_fx, '_0')
svr_fb_lst <- paste0(st_bisec_prefix, st_upper_fx, '_0')
svr_fxvr_name <- paste0('f', svr_root_x)
ls_pmap_vars <- unique(c(ls_svr_df_in_func, svr_root_x))

# B. Add initial a and b
df_bisec <- df %>% mutate(!!sym(svr_a_lst) := fl_lower_x, !!sym(svr_b_lst) := fl_upper_x)

# C. Evaluate function f(a_0) and f(b_0)
# 1. set x = a_0
# 2. evaluate f(a_0)
# 3. set x = b_0
# 4. evaluate f(b_0)
df_bisec <- df_bisec %>% mutate(!!sym(svr_root_x) := !!sym(svr_a_lst))
df_bisec <- df_bisec %>% mutate(
                !!sym(svr_fa_lst) :=
                  unlist(
                    pmap(df_bisec %>% select(ls_pmap_vars), fc_withroot)
                    )
                )
df_bisec <- df_bisec %>% mutate(!!sym(svr_root_x) := !!sym(svr_b_lst))
df_bisec <- df_bisec %>% mutate(
                !!sym(svr_fb_lst) :=
                  unlist(
                    pmap(df_bisec %>% select(ls_pmap_vars), fc_withroot)
                    )
                )

# D. Iteration Convergence Criteria
# fl_p_dist2zr = distance to zero to initalize
fl_p_dist2zr <- 1000
it_cur <- 0
while (it_cur <= it_iter_tol && fl_p_dist2zr >= fl_zero_tol ) {

  it_cur <- it_cur + 1

  # New Variables
  svr_a_cur <- paste0(st_bisec_prefix, st_lower_x, '_', it_cur)
  svr_b_cur <- paste0(st_bisec_prefix, st_upper_x, '_', it_cur)
  svr_fa_cur <- paste0(st_bisec_prefix, st_lower_fx, '_', it_cur)
  svr_fb_cur <- paste0(st_bisec_prefix, st_upper_fx, '_', it_cur)

  # Evaluate function f(a_0) and f(b_0)
  # 1. generate p
  # 2. generate f_p
  # 3. generate f_p*f_a
  df_bisec <- df_bisec %>% mutate(!!sym(svr_root_x) := ((!!sym(svr_a_lst) + !!sym(svr_b_lst))/2))
  df_bisec <- df_bisec %>% mutate(
                  !!sym(svr_fxvr_name) :=
                    unlist(
                      pmap(df_bisec %>% select(ls_pmap_vars), fc_withroot)
                      )
                  ) %>%
                  mutate(f_p_t_f_a = !!sym(svr_fxvr_name)*!!sym(svr_fa_lst))
  
  # fl_p_dist2zr = sum(abs(p))
  fl_p_dist2zr <- mean(abs(df_bisec %>% pull(!!sym(svr_fxvr_name))))

  # Update a and b
  df_bisec <- df_bisec %>%
                  mutate(!!sym(svr_a_cur) :=
                           case_when(f_p_t_f_a < 0 ~ !!sym(svr_a_lst),
                                     TRUE ~ !!sym(svr_root_x))) %>%
                  mutate(!!sym(svr_b_cur) :=
                           case_when(f_p_t_f_a < 0 ~ !!sym(svr_root_x),
                                     TRUE ~ !!sym(svr_b_lst)))
  
  # Update f(a) and f(b)
  df_bisec <- df_bisec %>%
                  mutate(!!sym(svr_fa_cur) :=
                           case_when(f_p_t_f_a < 0 ~ !!sym(svr_fa_lst),
                                     TRUE ~ !!sym(svr_fxvr_name))) %>%
                  mutate(!!sym(svr_fb_cur) :=
                           case_when(f_p_t_f_a < 0 ~ !!sym(svr_fxvr_name),
                                     TRUE ~ !!sym(svr_fb_lst)))
  
  # Drop past record possibly
  if(!bl_keep_iter) {
    df_bisec <- df_bisec %>% select(-one_of(c(svr_a_lst, svr_b_lst, svr_fa_lst, svr_fb_lst)))
  }
  
  # Save from last
  svr_a_lst <- svr_a_cur
  svr_b_lst <- svr_b_cur
  svr_fa_lst <- svr_fa_cur
  svr_fb_lst <- svr_fb_cur

  # Summar current round
  message(paste0('it_cur:', it_cur, ', fl_p_dist2zr:', fl_p_dist2zr))
}


# return
return(df_bisec)

}
```

## Prepare Inputs to Invoke Bisection Function

Here we use the same function as earlier, but modify number of parameters:

```{r}
# 
df <- tb_states_choices
fc_withroot <- function(fl_A, fl_alpha, x){
  # Function has four inputs hardcoded in: ar_nN_A, ar_nN_alpha, fl_rho, fl_N_agg
  # Also renamed fl_N to x, indicating this is the x we are shifting to look for zero. 
  ar_p1_s1 = exp((fl_A - ar_nN_A)*fl_rho)
  ar_p1_s2 = (fl_alpha/ar_nN_alpha)
  ar_p1_s3 = (1/(ar_nN_alpha*fl_rho - 1))
  ar_p1 = (ar_p1_s1*ar_p1_s2)^ar_p1_s3
  ar_p2 = x^((fl_alpha*fl_rho-1)/(ar_nN_alpha*fl_rho-1))
  ar_overall = ar_p1*ar_p2
  fl_overall = fl_N_agg - sum(ar_overall)
  return(fl_overall)
}
fl_lower_x <- 0
fl_upper_x <- 200
ls_svr_df_in_func <- c('fl_A', 'fl_alpha')
svr_root_x <- 'x'
fl_zero_tol = 10^-6
```

Additionally, let's prepare a very simple linear test:

```{r}
ar_intercept = seq(-10, -1, length.out = it_N_child_cnt)
ar_slope = seq(0.1, 1, length.out = it_N_child_cnt)
df_lines <- as_tibble(cbind(ar_intercept, ar_slope)) %>% rowid_to_column(var='ID')
ar_st_col_names = c('ID','fl_int', 'fl_slope')
df_lines <- df_lines %>% rename_all(~c(ar_st_col_names))
fc_withroot_line <- function(fl_int, fl_slope, x){
  return(fl_int + fl_slope*x)
}
fl_lower_x_line <- 0
fl_upper_x_line <- 100000
ls_svr_df_in_func_line <- c('fl_int', 'fl_slope')
svr_root_x_line <- 'x'
fl_zero_tol = 10^-6
```

## Test Function 

Now we call the function to see if it works:

```{r}
df_bisec <- fv_opti_bisect_pmap_multi_test(df, fc_withroot, fl_lower_x,fl_upper_x,
                                           ls_svr_df_in_func, svr_root_x, bl_keep_iter = TRUE)

# Do individual inputs sum up to total allocations available
fl_allocated <- sum(df_bisec %>% pull(svr_root_x))
if (fl_N_agg == fl_allocated){
  message('Total Input Available = Sum of Optimal Individual Allocations')
} else {
  cat('fl_N_agg =', fl_N_agg, ', but, sum(df_bisec %>% pull(svr_root_x))=', fl_allocated)
}

# Show all 
df_bisec
```

Call the same function, but do not save history:

```{r}
df_bisec <- suppressMessages(fv_opti_bisect_pmap_multi_test(df, fc_withroot, fl_lower_x, fl_upper_x,
                                                            ls_svr_df_in_func, svr_root_x, bl_keep_iter = FALSE))
df_bisec %>% select(-one_of('f_p_t_f_a'))
```

Call the linear root function:

```{r}
df_bisec <- fv_opti_bisect_pmap_multi_test(df_lines, fc_withroot_line,
                                           fl_lower_x_line, fl_upper_x_line,
                                           ls_svr_df_in_func_line, svr_root_x_line, bl_keep_iter = FALSE)
df_bisec %>% select(-one_of('f_p_t_f_a'))
```

## Test with PrjOptiAlloc log results

Use *df_esti* from [*ffy_opt_dtgch_cbem4*](https://fanwangecon.github.io/PrjOptiAlloc/reference/df_opt_dtgch_cbem4.html) function of project *PrjOptiAlloc*. Select only the *fl_A* and *fl_alpha* variables, should look like *tb_states_choices*. The difference is that *tb_states_choices* values were made up, but the *alpha* and *A* values from *df_esti* are estimated. 

The function is the same as before: *fc_withroot*, in fact, this function is the solution to the log linear allocation from problem [PrjOptiAlloc](https://fanwangecon.github.io/PrjOptiAlloc/).

```{r}
# Load Package and Data
library(PrjOptiAlloc)
ls_opti_alpha_A <- ffy_opt_dtgch_cbem4()
df_esti <- ls_opti_alpha_A$df_esti

# Select only A and Alpha and Index 
# %>% mutate(A_log =(A_log))
df_dtgch_cbem4_bisec <- df_esti[1:2,] %>% select(indi.id, A_log, alpha_log) 
ar_st_col_names = c('INDI_ID', 'fl_A', 'fl_alpha')
df_dtgch_cbem4_bisec <- df_esti[1:2,] %>% select(A_log, alpha_log) 
ar_st_col_names = c('fl_A', 'fl_alpha')
df_dtgch_cbem4_bisec <- df_dtgch_cbem4_bisec %>% rename_all(~c(ar_st_col_names))

# Bisect and Solve
ls_svr_df_in_func <- c('fl_A', 'fl_alpha')
svr_root_x <- 'x'
df_bisec_dtgch_cbem4 <- fv_opti_bisect_pmap_multi_test(df_dtgch_cbem4_bisec, fc_withroot,
                                                       0, fl_upper_x,
                                                       ls_svr_df_in_func, svr_root_x,
                                                       bl_keep_iter = FALSE)

# Do individual inputs sum up to total allocations available
fl_allocated <- sum(df_bisec_dtgch_cbem4 %>% pull(svr_root_x))
if (fl_N_agg == fl_allocated){
  message('Total Input Available = Sum of Optimal Individual Allocations')
} else {
  cat('fl_N_agg =', fl_N_agg, ', but, sum(df_bisec_dtgch_cbem4 %>% pull(svr_root_x))=', fl_allocated)
}
  
df_bisec_dtgch_cbem4
```